---
title: "Causal Inference"
output:
  github_document: default
  html_document: default
  word_document: default
---

```{r echo = T, results = 'hide', message=F, warning=F}
library(readxl)
library(stargazer)
library(ggplot2)
library(MatchIt)
library(data.table)
library(tableone)
library(rbounds)
library(MESS)
library(dplyr)
```


### Unit of Analysis: Employees
### Treatment: Overtime (Y/N)
### Outcome of Interest: Attrition (Y/N)

>Note: Results of some tests in this markdown may vary by one or two decimal places when re-run as such techniques depend on the 'random value' provided by R interpreter. Always set seed manually to get reproducible results.

## Descriptive Stats
```{r}
hr <- read.csv('WA_Fn-UseC_-HR-Employee-Attrition.csv')
detach()
attach(hr)
str(hr)
```

##Idea
Since our data is cross-sectional data, the experimental method used in this case is Matching. The ideal comparison group is selected that matches the treatment group using either a comprehensive baseline survey or time invariant characteristics. The matches are selected on the basis of similarities in observed characteristics.This assumes no selection bias based on unobserved characteristics

Propensity score matching is used to match treated and untreated observations on the estimated probability of being treated (propensity score). 



**Steps followed in propensity Score Matching (PSM):**

1. Need representative and comparable data for both treatment and comparison groups
2. Used a logit estimator to estimate attrition as a function of observable characteristics
3. Used predicted values from estimator to generate propensity score for all treatment and comparison group members to evaluate matching


##Check power of test
To test whether we have enought data to conclude any significant causal relationship, I have performed a statistical test (power test) in R with the sample size of treatment as input. We find that with this data set we can reliably establish a causal relationship if our average treatment effect is at least 15.77%

```{r}
power_t_test(n=364,type=c("two.sample"),alternative="one.sided",power=0.8,sig.level=.1)
```

##Covariate balance before matching
Checking covariate balance before matching, the high P values like EnvironmentSatisfaction indicate that covariates are not properly balanced between treatmemnt and control. 

```{r}
tabUnmatched <- CreateTableOne(vars=c("ï..Age", "BusinessTravel", "DailyRate", "Department", "DistanceFromHome", "Education", "EducationField", "EnvironmentSatisfaction", "Gender", "HourlyRate", "JobInvolvement", "JobLevel", "JobRole", "JobSatisfaction", "MaritalStatus", "MonthlyIncome", "MonthlyRate", "NumCompaniesWorked", "PercentSalaryHike", "PerformanceRating", "RelationshipSatisfaction", "StockOptionLevel", "TotalWorkingYears", "TrainingTimesLastYear", "WorkLifeBalance", "YearsAtCompany", "YearsInCurrentRole", "YearsSinceLastPromotion", "YearsWithCurrManager"), strata="OverTime", test=TRUE,data=hr)
print(tabUnmatched, smd=TRUE)
```



## Calculate Propensity Score
Calcuate propensity score to evaluate treatment control balance
```{r}
hr$PS<-glm( OverTime~+ ï..Age + BusinessTravel + DailyRate + Department + DistanceFromHome + Education + EducationField + EnvironmentSatisfaction + Gender + HourlyRate + JobInvolvement + JobLevel + JobRole + JobSatisfaction + MaritalStatus + MonthlyIncome + MonthlyRate + NumCompaniesWorked  + PercentSalaryHike + PerformanceRating + RelationshipSatisfaction + StockOptionLevel + TotalWorkingYears + TrainingTimesLastYear + WorkLifeBalance + YearsAtCompany + YearsInCurrentRole + YearsSinceLastPromotion + YearsWithCurrManager, data=hr, family = "binomial")$fitted.values
```


## Plot propensity distribution before matching
The distribution of data with the propensity scores for treatment and control before matching technique is applied
```{r}
hr$attrCode<-ifelse(hr$Attrition == 'Yes', 1, 0)
hr$overcode<-ifelse(hr$OverTime == 'Yes', 1, 0)

ggplot(hr, aes(x = PS)) + 
  geom_histogram(color = "white") + 
  facet_wrap(~OverTime) +theme_bw() + coord_flip()


```


## Applyling Matching technique 
```{r}
Match <- matchit(overcode~+ ï..Age + BusinessTravel + DailyRate + Department + DistanceFromHome + Education + EducationField  +  EnvironmentSatisfaction + Gender + HourlyRate + JobInvolvement + JobLevel + JobRole + JobSatisfaction + MaritalStatus + MonthlyIncome + MonthlyRate + NumCompaniesWorked +  PercentSalaryHike + PerformanceRating + RelationshipSatisfaction  + StockOptionLevel + TotalWorkingYears + TrainingTimesLastYear + WorkLifeBalance + YearsAtCompany + YearsInCurrentRole + YearsSinceLastPromotion + YearsWithCurrManager, data = hr, method = 'nearest', distance="logit", caliper=0.006,replace=FALSE)

Matched.ids <- data.table(match.data(Match))$EmployeeNumber
hr.Matched <- hr[hr$EmployeeNumber %in% Matched.ids,]

```

##Covariate balance after matching
Checking covariate balance after matching, the low P values indicate that covariates are not balanced between treatmemnt and control. 
```{R}


tabMatched <- CreateTableOne(vars=c("ï..Age", "BusinessTravel", "DailyRate", "Department", "DistanceFromHome", "Education", "EducationField", "EnvironmentSatisfaction", "Gender", "HourlyRate", "JobInvolvement", "JobLevel", "JobRole", "JobSatisfaction", "MaritalStatus", "MonthlyIncome", "MonthlyRate", "NumCompaniesWorked", "PercentSalaryHike", "PerformanceRating", "RelationshipSatisfaction", "StockOptionLevel", "TotalWorkingYears", "TrainingTimesLastYear", "WorkLifeBalance", "YearsAtCompany", "YearsInCurrentRole", "YearsSinceLastPromotion", "YearsWithCurrManager"), strata="OverTime", test=TRUE,data=hr.Matched)
print(tabMatched, smd=TRUE)
```


##Propensity score after matching
The distribution of data with the propensity scores for treatment and control after matching technique is applied, looks lot more balanced. 
```{r}
hr.Matched$PS<-glm( OverTime~+ ï..Age + BusinessTravel + DailyRate + Department + DistanceFromHome + Education + EducationField + EnvironmentSatisfaction + Gender + HourlyRate + JobInvolvement + JobLevel + JobRole + JobSatisfaction + MaritalStatus + MonthlyIncome + MonthlyRate + NumCompaniesWorked  + PercentSalaryHike + PerformanceRating + RelationshipSatisfaction + StockOptionLevel + TotalWorkingYears + TrainingTimesLastYear + WorkLifeBalance + YearsAtCompany + YearsInCurrentRole + YearsSinceLastPromotion + YearsWithCurrManager, data=hr.Matched, family = "binomial")$fitted.values

ggplot(hr.Matched, aes(x = PS)) + 
  geom_histogram(color = "white") + 
  facet_wrap(~OverTime) + coord_flip()+
   labs(x='Similarity', y='Distribution') + ggtitle('Data distribution of people working overtime')+theme(plot.title = element_text(hjust = 0.5))

```

##Evaluating affect of working overtime on attrition
Assuming there to be no relationship between working overtime and attrition, there is less than 0.01% chance for getting the sample statistic we got. This is within the (Type 1) error range we are willing to accept. 
Through this analysis we learn that working overtime increases probability of leaving the firm by 19.7%

```{r}
t.test(attrCode ~overcode, data=hr.Matched)
summary(lm(attrCode ~overcode , data = hr.Matched ))
```

##Test of heterogeneity (Interaction Effect)
Tests whether the relationship between overtime and attrition is different for different individual characteristics of the employees. Such heterogeneity has been evaluated using interaction terms

###Heterogeneity on relationship status
If we assume the relationship between working overtime and attrition doesn’t change with the marital status of an individual, there is 0.07% chance for getting the sample statistic we got. This is within the error range we are willing to accept. Hence, marital status has statistically significant effect on the relationship between working overtime and attrition.
If an individual is single and working overtime his/her propensity to leave the firm increases by 0.22 over a married individual. 

```{r}
summary(lm(attrCode ~overcode * MaritalStatus, data = hr.Matched ))
```

###Heterogeneity on stock option level
If we assume the relationship between working overtime and attrition doesn’t change with the stock option level of an individual, there is a 0.1% chance for getting the sample statistic we got. This is within the error range we are willing to accept.
Hence, stock option level has statistically significant effect on the relationship between working overtime and attrition. Individuals are more accepting to working overtime when most of their stock options are locked. However, as more and more stock options can be encashed the propensity to attrite also increases. For example: If an individual has locked stock options and working overtime his/her propensity to leave the firm decreases by 14% over an individual who doesn’t have stock options. 
```{r}
summary(lm(attrCode ~overcode * factor(StockOptionLevel), data = hr.Matched ))
```


###Heterogeneity on job role
If we assume the relationship between working overtime and attrition doesn’t change with the job level of an individual, there is a 0.01% chance for getting the sample statistic we got. This is within the error range we are willing to accept.
Hence, job level has statistically significant effect on the relationship between working overtime and attrition.
We observed that propensity to attrite when working overtime drop during the early career growth on an individual and starts to increase again after 3 pomostions. For example: If an individual had 3 promotions and is working overtime his/her propensity to leave the firm decreases by 31% had this been an individual’s entry level job.
This could be because young employees could me more aggressive for career growth and willing to work overtime, which fades once they are promoted by few levels.
```{r}
summary(lm(attrCode ~overcode * factor(JobLevel), data = hr.Matched ))
```


##Assumptions: 
Following are some assumptions made in this analysis

1. No Interference: Employee interactions can lead to over-worked employees directly or indirectly influencing other employees to quit. We assume that’s not happening significantly in this case.
2. Temporal precedence: Employees who quit can no longer work, we can be sure that working overtime always happens before attrition.
3. No endogeneity: We assume sample data we have is representative of population. No measurement error, especially in identifying overworking employees. And that all drivers are taken into consideration.
4. No Omitted variables: We assume that treatment only depends on observed confounders considered in the model.
5. Matching on outcomes: None of the employee characteristics should be influenced by treatment, this is a safe assumption to make for the variables of this dataset.
