---
title: "Descriptive Statistics"
output:
  github_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Loading the required R packages
```{r echo = T, results = 'hide', message=F, warning=F}
suppressPackageStartupMessages({
library(readxl)
library(stargazer)
library(MatchIt)
library(data.table)
library(tableone)
library(rbounds)
library(randomForest)
library(ggplot2)
library(ggthemes)
library(extrafont)
library(plyr)
library(scales)
library(dplyr)
library(tidyr)
library(keep)
})

```

#### Unit of Analysis: Employees
#### Treatment: Overtime (Y/N)
#### Outcome of Interest: Attrition (Y/N)


# Loading the dataset and summarising features
```{r}
hr <- read.csv('WA_Fn-UseC_-HR-Employee-Attrition.csv')
detach()
attach(hr)
#head(hr)
#colnames(hr)
#summary(hr)
str(hr)

```

# How many of the employees were treated / worked Overtime? 
```{r}
length(unique(hr$EmployeeNumber[hr$OverTime=='Yes']))
```


# How many of the employees did not work Overtime? 
```{r}
length(unique(hr$EmployeeNumber[hr$OverTime=='No']))
```


# What percentage of the employees received the treatment / worked overtime?
```{r}
prop_overtime <- length(unique(hr$EmployeeNumber[hr$OverTime=='Yes'])) / length(unique(hr$EmployeeNumber[hr$OverTime=='No']))
round(prop_overtime * 100, digits=2)
```


# Plot showing the count of Employees working Overtime vs Not Overtime
```{r}
ggplot(data=hr, aes(OverTime)) +
  theme_bw() +
  geom_bar(fill = "dark red") +
  ggtitle("Overtime (Yes / No)")
```



# Plot showing the count of Employees Attrited vs Not Attrited
```{r}
ggplot(data=hr, aes(Attrition)) +
  theme_bw() +
  geom_bar(fill = "dark red") +
  ggtitle("Attrition (Yes / No)")
```



# Count of Attrtition in employees who are working Overtime vs those who are not
```{r}
plt1 <- ggplot() +
  theme_bw() +
  geom_bar(aes(x = OverTime, y = Attrition, fill  = Attrition), data = hr, stat = "identity")
plt1
```



# Percentage of Attrtition in employees who are working Overtime vs those who are not
```{r}
hr_sample <- hr %>% group_by(OverTime, Attrition) %>% dplyr::summarise(check1 = n()) %>% group_by(OverTime) %>%   mutate(check2 = sum(check1)) %>% ungroup() %>% mutate(percentage = round((check1/check2)*100, digits = 2))

plt2 <- ggplot() +
  theme_bw() +
  geom_bar(aes(x = OverTime, y = percentage, fill  = Attrition), data = hr_sample, stat = "identity")
plt2

hr_sample <- ddply(hr_sample, .(OverTime), transform, pos = (0.5 * percentage))
fill <- c("#40b8d0", "#b2d183")
plt2 <- plt2 +
  geom_text(data=hr_sample, aes(x = OverTime, y = pos, label = paste0(percentage,"%")), size=4) +
  scale_fill_manual(values=fill) +
  xlab("Overtime (Yes / No)") + ylab("Employee Attrition (%)") +
  ggtitle("Attrition(%) in Overtime vs Non-Overtime Employees")
plt2
```



# Univariate Analysis

## Looking at the numeric variables
```{r}
ggplot(data=hr, aes(ï..Age)) + 
  theme_bw() +
  xlab("Age of Employees") + ylab("Number of Employees") +
  geom_histogram(fill = "dark red") +
  ggtitle("Age Distribution")
```



## Distribution of Employees by Gender
```{r}
(length(unique(hr$EmployeeNumber[hr$Gender=='Male'])) / length(unique(hr$EmployeeNumber))) * 100
```



```{r}
ggplot(data=hr, aes(Gender)) + 
  theme_bw() +
  xlab("Gender of Employees") + ylab("Number of Employees") +
  geom_histogram(stat = 'count',fill = "dark red") +
  ggtitle("Gender Distribution")
```



```{r}
hr[c('DailyRate', 'DistanceFromHome', 'ï..Age', 'HourlyRate', 'MonthlyIncome','NumCompaniesWorked')] %>%
  select_if(is.numeric) %>%
  gather() %>%                             
  ggplot(aes(x = value)) +
  theme_bw() +
  facet_wrap(~ key, scales = "free") +  
  geom_histogram(fill = "dark red")
```



```{r}
hr[c('PercentSalaryHike', 'TotalWorkingYears', 'YearsAtCompany', 'YearsInCurrentRole', 'YearsSinceLastPromotion', 'YearsWithCurrManager')] %>%
  select_if(is.numeric) %>%
  gather() %>%                             
  ggplot(aes(x = value)) +
  theme_bw() +
  facet_wrap(~ key, scales = "free") +  
  geom_histogram(fill = "dark red")
```









